A Domain-Specific Optimizing Compiler\\for Solving Partial Differential Equations

### 2014-02-19 paper/bare_conf.tex
Refine padding/data alignment, start register tiling

% Padding and Data Alignment
Auto-vectorization of assembly code that computes $A$ can be less effective if data are not aligned and if the length of the innermost loop is small compared to the hardware vector length ($vl$). Data alignment is enforced in two steps. Initially, both arrays and matrices are allocated to addresses that are multiples of $vl$. Then, matrices are padded by rounding the number of columns to the nearest multiple of $vl$. For example, assume the original size of a matrix is 3$\times$3, and the underlying machine possesses AVX, which implies $vl=4$ since a vector register is 256 bits long and our kernels use double-precision floating-point values (64 bits). Then, the padded matrix on this architecture is 3$\times$4. Padding of all matrices involved in the evaluation of the stiffness matrix allows us to safely round the loop trip count to the nearest multiple of $vl$. This avoids the introduction of a remainder (scalar) loop from the compiler, which would be responsible for inefficient vectorization.

% Register tiling in its two forms
One notable problem assembly kernels are exposed to concerns register allocation and locality. There exist two critical situations that can negatively impact performance. The first occurs when trip counts are sufficiently large to prevent register reuse along the innermost dimension ($k$). A combination of the two scenarios is also possible. We want to exploit the small size of the iteration space and the structure of the computation to optimize register management.

>>>>

% Padding and Data Alignment
Auto-vectorization of assembly code that computes $A$ can be less effective if data are not aligned and if the length of the innermost loop is small compared to the hardware vector length ($vl$). Data alignment is enforced in two steps. Initially, both arrays and matrices are allocated to addresses that are multiples of $vl$. Then, matrices are padded by rounding the number of columns to the nearest multiple of $vl$. For example, assume the original size of a matrix is 3$\times$3, and the underlying machine possesses AVX, which implies $vl=4$ since a vector register is 256 bits long and our kernels use double-precision floating-point values (64 bits). Then, the padded matrix on this architecture is 3$\times$4. The compiler is explicitly informed about data alignment using a suitable pragma. Padding of all matrices involved in the evaluation of the stiffness matrix allows us to safely round the loop trip count to the nearest multiple of $vl$. This avoids the introduction of a remainder (scalar) loop from the compiler, which would be responsible for inefficient vectorization.

% Register tiling in its two forms
One notable problem assembly kernels are exposed to concerns register allocation and register locality. The critical situation occurs when loop trip counts and number of accessed matrices along the $j$ and $k$ dimensions are such that available vector registers are under-utilized. Due to a kernel's working set fitting the L1 cache, it is remarkably important to optimize register management in order to maximize performance. Canonical optimizations (i.e. non domain-driven), such as loop interchange, unroll, unroll-and-jam and register tiling are typically employed to deal with this problem. In our compiler we support these optimizations either by means of explicit source code transformation (interchange, register tiling, unroll-and-jam) or indirectly advising the compiler through standard pragmas (unroll). With register tiling, in particular, we slice the innermost loop into rectangular blocks of identical size (except the ``reminder'' block), and we apply unrolling to individual blocks.

%are sufficiently large to prevent register reuse along the innermost dimension $k$. A combination of the two scenarios is also possible. We want to exploit the small size of the iteration space and the structure of the computation to optimize register management.
